---
title: "Arctic Data Center Training"
author: "Natasha Griffin"
date: "10/19/2020"
output: html_document
---

Jeanette keeps this chunk in because it's easy to edit and/or reference to help format other sections.
For example, when echo = FALSE, the code will be run but not shown for the knitted document.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Hello **world!** 
## *this is Tasha's Rmarkdown*
Three facts: 

1. Microbes are cool
2. Subglacial lakes contain ~15% of all liquid but not a proportionate amount of carbon
3. The end

### New exercise
cmd+opt+i
```{r}
x <- 3*4
x
```

### Self-directed exercise
(replaced mine with Tatiana's because hers was cooler)
```{r}

t=seq(0,10,0.1)
y=sin(t)
plot(t,y,type="l", xlab="time", ylab="wave")

```

## Introduction
This dataset came from North Pole Environmental Observatory bottle chemistry data collected in 2009 by Craig Tweedie (ayyyy) and downloaded from the Arctic Data Center [here](https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X).

#Install packages
```{r}

library(readr)

```

# Read in data
Note: message=FALSE gets rid of the summary generated by read_csv
```{r, message=FALSE}

bg_chem <- read_csv("BGchem2008data.csv")

```

## Analysis
* Calculate summary statistics

```{r}

nitrate <- mean(bg_chem$NO3)
nitrite <- mean(bg_chem$NO2)
amm <- mean(bg_chem$NH4)
phos <- mean(bg_chem$P)

```


* Calculate mean Redfield ratio
* Make a plot
```{r}

mean_temp <- mean(bg_chem$CTD_Temperature)

```

The mean temperature was `r round(mean_temp)`.

```{r}

plot(bg_chem$CTD_Depth, bg_chem$CTD_Temperature)

```

